---
title: "K-Nearest Neighbour Analysis"
output: html_document
date: "2024-06-24"
---

# K-Nearest Neighbour Analysis

## Overview

## Code Implementation

### Load Libraries

```{r}
pacman::p_load(
  # Data manipulation
  tidyverse,
  
  #KNN:
  magrittr, rio, here, caret, kknn, performance, pROC, class, cluster,
  
  # Other:
  ggthemes, scale
)
```

### Import the Data

```{r}
df = read.csv("~/Library/CloudStorage/OneDrive-UniversidaddeCostaRica/GitHub/PrediccioÌn_Cancer/data/processed/data.csv")[,-c(1,2)]
head(df)
```

#### Transform to the correct datatypes

```{r}
df = df %>% 
  mutate_at(.vars = vars(Race, Education, 
                         Coverage, GeneralHealth, 
                         Children, Depression,
                         Exercise, Smoker,
                         Mammo, Breast), 
            .funs = as.factor)

df$Breast = relevel(x = df$Breast, ref = "No Breast Cancer")
```

#### Splitting the Dataset

We'll be applying 10-fold for cross validation

1.  Apply cross validation to see the optimal decision boundary
    1.  We'll apply the following decision boundaries to compare [0.2, 0.3, 0.4, 0.5, 0.6, 0.7]. Where if the y_predicted is bigger than the respective decision boundary will predict 1 \~ Breast Cancer

```{r}
folds = function(df, k, seed){
  
  set.seed(seed = seed)
  
  fold_group = sample(rep(1:k, length.out = nrow(df)))
  
  df$fold = fold_group
  
  return(df)
  
}
```

```{r}
df_folds = folds(df, k = 10, seed = 8)
```


### Implementing K Nearest Neighbour

```{r}
knn_model = function(df_folds, n_test_fold, k){
  
  predictions = c()
  
  train_len = nrow(df_folds[df_folds$fold != n_test_fold,-14])
  train_rows = rownames(df_folds[df_folds$fold != n_test_fold,-14])
  
  test_len = nrow(df_folds[df_folds$fold == n_test_fold,-14])
  
  distances = as.matrix(daisy(rbind(df_folds[df_folds$fold != n_test_fold,-c(14,ncol(df_folds))],
                                     df_folds[df_folds$fold == n_test_fold,-c(14,ncol(df_folds))])))
  
  distances = distances[1:train_len,-c(1:train_len)]
  
  distances_order = apply(distances, 2, order)
  
  distances_order = as.matrix(distances_order)
  
  rownames(distances_order) = train_rows
  
  for (test_value in 1:test_len){
   
    nearest_neighbours = df_folds[head(order(distances[,test_value]), k), 14]
    
    predictions[test_value] = names(which.max(table(nearest_neighbours)))
    
    predictions[test_value] = ifelse(predictions[test_value] == "Breast Cancer", 1, 0)
    
  }
  
  return(factor(predictions, levels = c(0, 1), labels = c("No Breast Cancer", "Breast Cancer")))
}
```

```{r}
performance_knn = function(predicted, y_test){
  
  
    cm = table(predicted, y_test)
    
    results = c(sum(diag(cm)) / sum(cm), #Accuracy
                1 - sum(diag(cm)) / sum(cm), #Precision
                cm[2,2] / sum(cm[,2]), #Sensitivity
                cm[1,1] / sum(cm[,1]), #Specificity
                cm[2,1] / (cm[2,1]+cm[1,1]) , #False Positives,
                cm[1,2]/ (cm[1,2]+cm[2,2]) , #False Positives,
                (sum(diag(cm)) / sum(cm) - sum(rowSums(cm) * colSums(cm)) / sum(cm)^2) / (1 - sum(rowSums(cm) * colSums(cm)) / sum(cm)^2), #Kappa
                auc(roc(y_test, as.numeric(predicted))) #AUC
    )
    
    return(results)
  
}
```

```{r}
knn_validation = function(df_folds, k){
  
  final_df = matrix(data = NA,
                  nrow = max(df_folds$fold), ncol = 10)
  
  for (fold in 1:max(df_folds$fold)){
    
    predictions = knn_model(df_folds = df_folds, n_test_fold = fold, k = k)
    
    results = performance_knn(predicted = predictions, y_test = df_folds[df_folds$fold == fold, 'Breast'])
   
    results = c(fold, results, k)
    
    final_df[fold,] = results  
  }
  
  final_df = as.data.frame(final_df)
  colnames(final_df) = c('Test Fold', "Accuracy", "Precision",
                         "Sensitivity", "Specificity", "Fake_Positives",
                         "Fake_Negatives", "Kappa", "auc", 'k')
  
  
  return(final_df)
}
```


```{r}
prueba = c(3,7,13,21,29)
promedios = matrix(data = NA, ncol = 4, nrow = length(prueba))
sd_matriz = matrix(data = NA, ncol = 4, nrow = length(prueba))

for (i in 1:length(prueba)){
  
  knn_base = knn_validation(df_folds = df_folds, k = prueba[i])
  
  promedios[i,1] = suppressMessages(round(mean(knn_base$Accuracy)*100,1))
  promedios[i,2] = suppressMessages(round(mean(knn_base$Precision)*100,1))
  promedios[i,3] = suppressMessages(round(mean(knn_base$Fake_Negatives)*100,1))
  promedios[i,4] = suppressMessages(round(mean(knn_base$auc)*100,1))
  
  sd_matriz[i,1] = suppressMessages(round(sd(knn_base$Accuracy)*100,1))
  sd_matriz[i,2] = suppressMessages(round(sd(knn_base$Precision)*100,1))
  sd_matriz[i,3] = suppressMessages(round(sd(knn_base$Fake_Negatives)*100,1))
  sd_matriz[i,4] = suppressMessages(round(sd(knn_base$auc)*100,1))
  
}
```
```{r}
df_promedios = as.data.frame(promedios)

colnames(df_promedios) = c("Precision", "Error", "FalsosNegativos", "AUC")
rownames(df_promedios) = prueba

df_promedios
```

```{r}
df_sd = as.data.frame(sd_matriz)

colnames(df_sd) = c("Precision", "Error", "FalsosNegativos", "AUC")
rownames(df_sd) = prueba

df_sd
```


