---
title: "Logistic Regression Analysis"
output: html_document
date: "2024-06-23"
---

# Logistic Regression Analysis

## Overview

## Code Implementation

### Load Libraries

```{r}
pacman::p_load(
  # Data manipulation
  tidyverse,
  
  #Logistic Regression:
  magrittr, rio, here, caret, glmnet, performance, pROC, ROCR,
  
  # Other:
  ggthemes, scale
)
```

### Import the Data

```{r}
df = read.csv("~/Library/CloudStorage/OneDrive-UniversidaddeCostaRica/GitHub/PrediccioÌn_Cancer/data/processed/data.csv")[,-c(1,2)]
head(df)
```

#### Transform to the correct datatypes

```{r}
df = df %>% 
  mutate_at(.vars = vars(Race, Education, 
                         Coverage, GeneralHealth, 
                         Children, Depression,
                         Exercise, Smoker,
                         Mammo, Breast), 
            .funs = as.factor)

df$Breast = relevel(x = df$Breast, ref = "No Breast Cancer")
```

#### Splitting the Dataset

We'll be applying 10-fold for cross validation

1.  Apply cross validation to see the optimal decision boundary
    1.  We'll apply the following decision boundaries to compare [0.2, 0.3, 0.4, 0.5, 0.6, 0.7]. Where if the y_predicted is bigger than the respective decision boundary will predict 1 \~ Breast Cancer

```{r}
folds = function(df, k, seed){
  
  set.seed(seed = seed)
  
  fold_group = sample(rep(1:k, length.out = nrow(df)))
  
  df$fold = fold_group
  
  return(df)
  
}
```

```{r}
df_folds = folds(df, k = 10, seed = 8)
```

### Implementing Logistic Regression

```{r}
logistic_model = function(df_folds, n_test_fold, decision_boundary){
  
  logistic_model = glm(data = df_folds[df_folds$fold != n_test_fold,-ncol(df_folds)],
             formula = Breast ~ ., family = binomial)
  
  y_pred_prob = predict(logistic_model, 
                        type = 'response', 
                        newdata = df_folds[df_folds$fold == n_test_fold,-ncol(df_folds)])
  
  y_pred = ifelse(y_pred_prob > decision_boundary, 1, 0)
  
  predictions = matrix(data = NA, nrow = length(y_pred), ncol = 2)
  predictions = as.data.frame(predictions)
  predictions[,1] = y_pred_prob
  predictions[,2] = factor(y_pred, levels = c(0, 1), 
                           labels = c("No Breast Cancer", "Breast Cancer"))
  
  return(predictions)
  
}
```


```{r}
performance_logistic = function(y_pred, y_test, y_pred_prob){
  
  cm = confusionMatrix(y_test, y_pred ,positive = "Breast Cancer")
  
  results = c(
    cm$overall['Accuracy'],
    1 - cm$overall['Accuracy'], 
    cm$byClass['Sensitivity'], 
    cm$byClass['Specificity'],
    fp = cm$table[1,2]/(cm$table[1,2] + cm$table[1,1]), 
    fn = cm$table[2,1]/(cm$table[2,2] + cm$table[2,1]), 
    cm$overall['Kappa'],
    auc(roc(as.numeric(y_test), as.numeric(y_pred)))
  )
  
  return(results)
  
}
```

```{r}
logistic_validation = function(df_folds, decision_boundary){
  
  final_df = matrix(data = NA,
                    nrow = max(df_folds$fold), ncol = 10)
  
  for (fold in 1:max(df_folds$fold)){
    
    predictions = logistic_model(df_folds = df_folds, 
                            n_test_fold = fold, 
                            decision_boundary = decision_boundary)
    
    results = performance_logistic(y_pred = predictions[,2], 
                                   y_test = df_folds$Breast[df_folds$fold == fold],
                                   y_pred_prob = predictions[,1])
    
    results = c(fold, results, decision_boundary)
    
    final_df[fold,] = results
  
  }
  
  final_df = as.data.frame(final_df)
  colnames(final_df) = c('Test Fold', "Accuracy", "Precision",
                         "Sensitivity", "Specificity", "Fake_Positives",
                         "Fake_Negatives", "Kappa", "auc", 'decision_boundary')
  
  
  return(final_df)
  
}
```

```{r}
# for (i in c(0.2, 0.3, 0.4, 0.5, 0.6, 0.7)){
#  print(suppressMessages((logistic_validation(df_folds = df_folds, i)$Accuracy)))
# }
```

```{r}
prueba = c(0.2, 0.3, 0.5, 0.7, 0.8)
promedios = matrix(data = NA, ncol = 4, nrow = length(prueba))
sd_matriz = matrix(data = NA, ncol = 4, nrow = length(prueba))

for (i in 1:length(prueba)){
  
  promedios[i,1] = suppressMessages(round(mean(logistic_validation(df_folds = df_folds, decision_boundary = prueba[i])$Accuracy)*100,1))
  promedios[i,2] = suppressMessages(round(mean(logistic_validation(df_folds = df_folds, decision_boundary = prueba[i])$Precision)*100,1))
  promedios[i,3] = suppressMessages(round(mean(logistic_validation(df_folds = df_folds, decision_boundary = prueba[i])$Fake_Negatives)*100,1))
  promedios[i,4] = suppressMessages(round(mean(logistic_validation(df_folds = df_folds, decision_boundary = prueba[i])$auc)*100,1))
  
  sd_matriz[i,1] = suppressMessages(round(sd(logistic_validation(df_folds = df_folds, decision_boundary = prueba[i])$Accuracy)*100,1))
  sd_matriz[i,2] = suppressMessages(round(sd(logistic_validation(df_folds = df_folds, decision_boundary = prueba[i])$Precision)*100,1))
  sd_matriz[i,3] = suppressMessages(round(sd(logistic_validation(df_folds = df_folds, decision_boundary = prueba[i])$Fake_Negatives)*100,1))
  sd_matriz[i,4] = suppressMessages(round(sd(logistic_validation(df_folds = df_folds, decision_boundary = prueba[i])$auc)*100,1))
  
}
```



```{r}
df_promedios = as.data.frame(promedios)

colnames(df_promedios) = c("Precision", "Error", "FalsosNegativos", "AUC")
rownames(df_promedios) = prueba

df_promedios
```

```{r}
df_sd = as.data.frame(sd_matriz)

colnames(df_sd) = c("Precision", "Error", "FalsosNegativos", "AUC")
rownames(df_sd) = prueba

df_sd
```

