---
title: "Untitled"
output: html_document
date: "2024-06-26"
---

# Methods of Tree

## Overview

## Code Implementation

```{r}
pacman::p_load(
  # Data manipulation
  tidyverse,
  
  #Tree:
  magrittr, rio, here, caret, kknn, performance, pROC, class, rpart, rattle, DT,
  
  # Other:
  ggthemes, scale
)
```

### Import the Data

```{r}
df = read.csv("~/Library/CloudStorage/OneDrive-UniversidaddeCostaRica/GitHub/PrediccioÌn_Cancer/data/processed/data.csv")[,-c(1,2)]
head(df)
```

#### Transform to the correct datatypes

```{r}
df = df %>% 
  mutate_at(.vars = vars(Race, Education, 
                         Coverage, GeneralHealth, 
                         Children, Depression,
                         Exercise, Smoker,
                         Mammo, Breast), 
            .funs = as.factor)

df$Breast = relevel(x = df$Breast, ref = "No Breast Cancer")
```

#### Splitting the Dataset

We'll be applying 10-fold for cross validation

1.  Apply cross validation to see the optimal decision boundary
    1.  We'll apply the following hyperparameters to compare . Where if the y_predicted is bigger than the respective decision boundary will predict 1 \~ Breast Cancer

```{r}
folds = function(df, k, seed){
  
  set.seed(seed = seed)
  
  fold_group = sample(rep(1:k, length.out = nrow(df)))
  
  df$fold = fold_group
  
  return(df)
  
}
```

```{r}
df_folds = folds(df, k = 10, seed = 8)
```

### Implementing Tree Method

```{r}
tree_model = function(df_folds, n_test_fold, cp, maxdepth = 10){
  
  if (maxdepth == 10){
    
    tree = rpart(data = df_folds[df_folds$fold != n_test_fold,-ncol(df_folds)],
             formula = Breast ~., method = "class", 
             control=rpart.control(minsplit= 0.01 * nrow(df_folds[df_folds$fold !=n_test_fold,-ncol(df_folds)]),
                                   minbucket=0.005 * nrow(df_folds[df_folds$fold !=n_test_fold,-ncol(df_folds)]), 
                                   cp=cp, maxdepth = 10))
    
  }
  else{
    
    tree = rpart(data = df_folds[df_folds$fold != n_test_fold,-ncol(df_folds)],
             formula = Breast ~., method = "class", 
             control=rpart.control(minsplit= 0.01 * nrow(df_folds[df_folds$fold !=n_test_fold,-ncol(df_folds)]),
                                   minbucket=0.005 * nrow(df_folds[df_folds$fold !=n_test_fold,-ncol(df_folds)]), 
                                   cp=cp,
                                   maxdepth = maxdepth))
    
  }


  
  predictions = predict(tree, newdata =  df_folds[df_folds$fold == n_test_fold,],
                        type = "class")
  
  predictions = ifelse(predictions=="No Breast Cancer", yes = 0, no = 1)
  
  return(factor(predictions, levels = c(0, 1), labels = c("No Breast Cancer", "Breast Cancer")))

  
}
```

```{r}
performance_logistic = function(y_pred, y_test){
  
  cm = confusionMatrix(y_test, y_pred ,positive = "Breast Cancer")
  
  results = c(
    cm$overall['Accuracy'],
    1 - cm$overall['Accuracy'], 
    cm$byClass['Sensitivity'], 
    cm$byClass['Specificity'],
    fp = cm$table[1,2]/(cm$table[1,2]+cm$table[1,1]), 
    fn = cm$table[2,1]/(cm$table[2,1]+cm$table[2,2]), 
    cm$overall['Kappa'],
    auc(roc(as.numeric(y_test), as.numeric(y_pred)))
  )
  
  return(results)
  
}
```

```{r}
tree_validation = function(df_folds, cp, maxdepth = 10){
  
  final_df = matrix(data = NA,
                    nrow = max(df_folds$fold), ncol = 11)
  
  for (fold in 1:max(df_folds$fold)){
    
    predictions = tree_model(df_folds = df_folds, 
                            n_test_fold = fold, 
                            cp=cp,
                            maxdepth = maxdepth)
    
    results = performance_logistic(y_pred = predictions, 
                                   y_test = df_folds$Breast[df_folds$fold == fold])
    
    results = c(fold, results, cp, maxdepth)
    
    final_df[fold,] = results
  
  }
  
  final_df = as.data.frame(final_df)
  colnames(final_df) = c('Test Fold', "Accuracy", "Precision",
                         "Sensitivity", "Specificity", "Fake_Positives",
                         "Fake_Negatives", "Kappa", "auc", 'cp', 'maxdepth')
  
  
  return(final_df)
  
}
```

Calibrating cp

```{r}
prueba = c(0.01,0.075,0.005,0.0025,0.001)
promedios = matrix(data = NA, ncol = 4, nrow = length(prueba))
sd_matriz = matrix(data = NA, ncol = 4, nrow = length(prueba))

for (i in 1:length(prueba)){
  
  tree_base = suppressMessages(tree_validation(df_folds = df_folds, cp = prueba[i]))
  
  promedios[i,1] = suppressMessages(round(mean(tree_base$Accuracy)*100,1))
  promedios[i,2] = suppressMessages(round(mean(tree_base$Precision)*100,1))
  promedios[i,3] = suppressMessages(round(mean(tree_base$Fake_Negatives)*100,1))
  promedios[i,4] = suppressMessages(round(mean(tree_base$auc)*100,1))
  
  sd_matriz[i,1] = suppressMessages(round(sd(tree_base$Accuracy)*100,1))
  sd_matriz[i,2] = suppressMessages(round(sd(tree_base$Precision)*100,1))
  sd_matriz[i,3] = suppressMessages(round(sd(tree_base$Fake_Negatives)*100,1))
  sd_matriz[i,4] = suppressMessages(round(sd(tree_base$auc)*100,1))
  
}
```
```{r}
df_promedios = as.data.frame(promedios)

colnames(df_promedios) = c("Precision", "Error", "FalsosNegativos", "AUC")
rownames(df_promedios) = prueba

df_promedios
```

```{r}
df_sd = as.data.frame(sd_matriz)

colnames(df_sd) = c("Precision", "Error", "FalsosNegativos", "AUC")
rownames(df_sd) = prueba

df_sd
```

Calibrating maxdepth

```{r}
prueba = c(24,17,10,7,4)
promedios = matrix(data = NA, ncol = 4, nrow = length(prueba))
sd_matriz = matrix(data = NA, ncol = 4, nrow = length(prueba))

for (i in 1:length(prueba)){
  
  tree_base = suppressMessages(tree_validation(df_folds = df_folds, cp = 0.001, maxdepth = prueba[i]))
  
  promedios[i,1] = suppressMessages(round(mean(tree_base$Accuracy)*100,1))
  promedios[i,2] = suppressMessages(round(mean(tree_base$Precision)*100,1))
  promedios[i,3] = suppressMessages(round(mean(tree_base$Fake_Negatives)*100,1))
  promedios[i,4] = suppressMessages(round(mean(tree_base$auc)*100,1))
  
  sd_matriz[i,1] = suppressMessages(round(sd(tree_base$Accuracy)*100,1))
  sd_matriz[i,2] = suppressMessages(round(sd(tree_base$Precision)*100,1))
  sd_matriz[i,3] = suppressMessages(round(sd(tree_base$Fake_Negatives)*100,1))
  sd_matriz[i,4] = suppressMessages(round(sd(tree_base$auc)*100,1))
  
}
```

```{r}
df_promedios = as.data.frame(promedios)

colnames(df_promedios) = c("Precision", "Error", "FalsosNegativos", "AUC")
rownames(df_promedios) = prueba

df_promedios
```

```{r}
df_sd = as.data.frame(sd_matriz)

colnames(df_sd) = c("Precision", "Error", "FalsosNegativos", "AUC")
rownames(df_sd) = prueba

df_sd
```

